Converting Neural Networks to Gaussian Processes for Sequential Decision Making Under Uncertainty

### 2023-04-18 paper/main.tex
updated title

%\title{Investigatin Uncertainty Quantification in Model-based Reinforcement Learning}
% \title{Model-based Reinforcement Learning with Fast Posterior Updates}
%\title{Sequential Decision-Making under Uncertainty with Big Data}
\title{Neural Network to Vatiational Sparse Gaussian Process: For Adaptive Exploration}

@@

\begin{abstract}
% Building systems that adapt to new information and express uncertainty is essential to deploy methods in real-world applications e.g. healthcare, autonomous vehicles. While gradient-based training is symbiotic to modern deep learning, it is rigid regarding adding further information, usually involving retraining from scratch. Converting a trained neural network to its dual parameterization we can view the parameters of the model as data dependent terms instead of the weights. Furthermore, we can then build a sparse model to form a variational dual sparse Gaussian process. The resultant sparse GP has two obvious advantages: scalable principled uncertainty estimation and the ability adapt to new data without gradient based retraining. Our proposed method can be deployed in any sequential decision-making process, where new information is received and retraining from scratch is not available. We show its effectiveness in model-based reinforcement learning using updated information for more efficient exploration.
%
  Building learning-based systems that can adapt to new data and express uncertainty is essential for their deployment in real-world applications, e.g. healthcare and robotics. However, modern deep learning cannot adapt to new data because it relies upon gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data without further training. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP). Importantly, our method 1) enables fast adaptation to new data, 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets. Essentially, our method reaps the benefits of nonparametric models without their well-known issues. We demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning and we demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets.

>>>>

%\title{Investigatin Uncertainty Quantification in Model-based Reinforcement Learning}
% \title{Model-based Reinforcement Learning with Fast Posterior Updates}
%\title{Sequential Decision-Making under Uncertainty with Big Data}
% \title{Neural Network to Vatiational Sparse Gaussian Process: For Adaptive Exploration}
% \title{Neural Network to Sparse Variational Gaussian Process: For Updates in Sequential Decision Making}
% \title{Adapting Neural Networks to New Data For Updates in Sequential Decision Making via Gaussian Processes}
\title{Converting Neural Networks to Gaussian Processes for Sequential Decision Making Under Uncertainty}

@@

\begin{abstract}
% Building systems that adapt to new information and express uncertainty is essential to deploy methods in real-world applications e.g. healthcare, autonomous vehicles. While gradient-based training is symbiotic to modern deep learning, it is rigid regarding adding further information, usually involving retraining from scratch. Converting a trained neural network to its dual parameterization we can view the parameters of the model as data dependent terms instead of the weights. Furthermore, we can then build a sparse model to form a variational dual sparse Gaussian process. The resultant sparse GP has two obvious advantages: scalable principled uncertainty estimation and the ability adapt to new data without gradient based retraining. Our proposed method can be deployed in any sequential decision-making process, where new information is received and retraining from scratch is not available. We show its effectiveness in model-based reinforcement learning using updated information for more efficient exploration.
%
  Building learning-based systems that can adapt to new data and express uncertainty is essential for their deployment in real-world applications, e.g. healthcare and robotics. However, modern deep learning methods cannot adapt to new data because they rely on gradient-based optimisation, which requires retraining with the addition of new data. In contrast, nonparametric approaches -- such as Gaussian processes -- can adapt to new data. However, they are limited to low-dimensional and small data sets. We present a method that enables any trained neural network (NN) to be converted to a sparse variational Gaussian process (SVGP). Importantly, our method 1) enables fast adaptation to new data, 2) provides well-calibrated uncertainty estimates, and 3) scales to high-dimensional and large data sets. Essentially, our method reaps the benefits of nonparametric models without their well-known issues. We demonstrate the effectiveness of our model's fast adaptation by showing that it can improve sample-efficiency in model-based reinforcement learning and we demonstrate our state-of-the-art NN uncertainty quantification with experiments using UCI data sets.
